---
title: "Homework 2"
author: "Shina Min"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

We clean the 538 `pols` data, which provides information on the number of national politicians who are democratic or republican at any given time. There are some values for which `prez_gop` is `2` -- these are months in which Ford became President following Nixon's resignation. In the new `president` variable created as part of our data cleaning, we code these as `gop` (same as values when `prez_gop` is `1`).

```{r clean_538_pols}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols =
  read_csv("./fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez"))
```

We also clean the 538 `snp` data, which contains information related to Standard & Poor’s stock market index.

```{r clean_538_snp}
snp = 
  read_csv(
    "./fivethirtyeight_datasets/snp.csv",
    col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```

Finally, we tidy the `unemployment` data so that it can be merged with the `pols` and `snp` datasets.

```{r clean_538_unemp}
unemployment = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Now we merge the three datasets!

```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```

Notice that there are some `NA` values in the `close` and `unemployment` variables, which indicate that the value of these variables is missing at those locations.

Let's talk about the 538 datasets. The `pols` data has `r nrow(pols)` observations and `r ncol(pols)` variables and tells us about the party affiliation distribution (democrat or republican) for governors and senators for a given year from years `r pols |> pull(year) |> min()` to `r pols |> pull(year) |> max()`. It also tells us whether the sitting president was a democrat or republican. The `snp` data has `r nrow(snp)` observations and `r ncol(snp)` variables, ranging from years `r snp |> pull(year) |> min()` to `r snp |> pull(year) |> max()`. The `unemployment` data has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables ranging from years `r unemployment |> pull(year) |> min()` to `r unemployment |> pull(year) |> max()`. In Januarys in or after 1975 in which a democrat was president, the **average unemployment rate was `r filter(data_538, month == "January", year >= 1975, president == "dem") |> pull(unemployment) |> mean() |> round(2)`**.  The average unemployment rate over the same time period in which a republican was president was `r filter(data_538, month == "January", year >= 1975, president == "gop") |> pull(unemployment) |> mean() |> round(2)`.


### Problem 2

This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website.

Read and clean the Mr. Trash Wheel sheet:

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Reading and cleaning Trash Wheel Collection Data

```{r}
mr_trash_wheel =
  read_excel("./202309 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel_data = 1,
         year = as.numeric(year))

prof_trash_wheel =
  read_excel("./202309 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             range = cell_cols("A:M")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel_data = 2)

gwy_trash_wheel =
  read_excel("./202309 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda Trash Wheel",
             range = cell_cols("A:J")) %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel_data = 3)
```

## Merging three data sets

```{r}
trash_wheel_data =
 bind_rows(mr_trash_wheel, prof_trash_wheel, gwy_trash_wheel) %>%
  janitor::clean_names()
```

In the `mr_trash_wheel` dataset, there are 15 variables, and 584 observations. The average number of homes powered is 53.3 and the total weight of trash collected is 1875 tons.

The `prof_trash_wheel` dataset includes 14 variables and 106 observations. The average number of homes powered is 34. The total weight of trash collected is 216 tons.

Lastly, the `gwy_trash_wheel` dataset has 13 variables and 155 observations. The average number of homes powered is 48.6 and the total weight of trash collected is 452 tons. The total number of cigarette butts collected by Gwynnda in July of 2021 is 16300.


### Problem 3

Import, clean, and tidy the dataset of baseline demographics. Ensure that sex and APOE4 carrier status are appropriate encoded (i.e. not numeric), and remove any participants who do not meet the stated inclusion criteria (i.e. no MCI at baseline). Discuss important steps in the import process and relevant features of the dataset. How many participants were recruited, and of these how many develop MCI? What is the average baseline age? What proportion of women in the study are APOE4 carriers?

```{r}
mci_df =
  read_csv("data_mci/MCI_baseline.csv", skip = 1) %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `0` = "male", `1` = "female")) %>%


    
  
  

```


Similarly, import, clean, and tidy the dataset of longitudinally observed biomarker values; comment on the steps on the import process and the features of the dataset.



Check whether some participants appear in only the baseline or amyloid datasets, and comment on your findings. Combine the demographic and biomarker datasets so that only participants who appear in both datasets are retained, and briefly describe the resulting dataset; export the result as a CSV to your data directory.

